Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period

### 2020-04-14 analysis/paper/paper.Rmd
edits on personal ornaments section based on BF's comment

Ornaments (Figure \@ref(fig:ornaments-image)) were found in a variety of archaeological contexts including post-holes area, burials, and middens. This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)). Occupation floors were not identified during excavation. We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site. There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan. Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173. The high number of ornaments in these burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007]. We exclude burials from this analysis because most of them date to the pre-European period (n = 26), limiting the usefulness of comparisons between the periods. In addition to `r sam_post_hole` ornaments from dwelling place contexts and 3,173 from burial contexts, there are 27 ornaments found in midden contexts. We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative on social inequality indicated by uneven distributions of ornaments.

```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class. A: carnelian beads, B: bells, C: glass beads and gold-foil beads, D: metal rings. Photographs are presented in the same order as those subtypes in the table but from left to right instead. The photographs of B, C, D classes were from original excavation report (Chen 2007)", echo=FALSE}
knitr::include_graphics(here("analysis", "figures", "ornament.png"))
```

@@

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "small bead", "small (0.5-1 cm)", Type),
         Type = ifelse(Type == "glass bead", "medium (1-2 cm)", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type),
         Type = ifelse(Type == "large thin string", "thin large", Type),
         Categories = ifelse(Categories == "Agate bead", "Carnelian bead", Categories),
         Categories = ifelse(Categories == "Golden bead", "Gold-foil bead", Categories),
         Categories = ifelse(Categories == "Bell", "Metal bell", Categories))
ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>%
  filter(!Type %in% c("other",
                      "unclassified")) %>% 
  arrange(Categories, desc(`European Presence`))
# making table
knitr::kable(ornaments_shape,
             caption = "Ornament subtype at Kiwulan. The numbers represent artifact counts ")

>>>>

Ornaments (Figure \@ref(fig:ornaments-image)) were found in a variety of archaeological contexts including post-holes area, burials, and middens. This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)). Occupation floors were not identified during excavation. We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site. There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan. Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173. The high number of ornaments in these burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007]. We exclude burials from this analysis because most of them date to the pre-European period (n = 26), limiting the usefulness of comparisons between the periods. In addition to `r sam_post_hole` ornaments from dwelling place contexts and 3,173 from burial contexts, there are 27 ornaments found in midden contexts. We focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments.

```{r, ornaments-image, fig.cap = "Subtypes of ornament in each major class. A: carnelian beads, B: bells, C: glass beads and gold-foil beads (far right), D: metal rings. Photographs are presented in the same order as those subtypes in the table but from left to right instead. The photographs of B, C, D classes were from original excavation report (Chen 2007)", echo=FALSE}
knitr::include_graphics(here("analysis", "figures", "ornament.png"))
```

@@

```{r tidy-data-for-table}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "small bead", "small (0.5-1 cm)", Type),
         Type = ifelse(Type == "glass bead", "medium (1-2 cm)", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type),
         Type = ifelse(Type == "large thin string", "thin large", Type),
         Categories = ifelse(Categories == "Agate bead", "Carnelian bead", Categories),
         Categories = ifelse(Categories == "Golden bead", "Gold-foil bead", Categories),
         Categories = ifelse(Categories == "Bell", "Metal bell", Categories))
ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring")) %>%
  filter(!Type %in% c("other",
                      "unclassified")) %>% 
  arrange(Categories, desc(`European Presence`))
# making table
knitr::kable(ornaments_shape,
             caption = "Frequencies of ornaments by subtype at Kiwulan. Frequency represents artifact counts ")
